---
title: "Strawberry Exploration"
format: html
editor: visual
---

## Preliminaries

```{r}
## if necessary, import libraries
library(dplyr)
```

```{r}
strawberry <- read.csv("straw1.csv")
```

```{r}
glimpse(strawberry)
```

drop_one_value_col(df): drop a column that only contains one value

```{r}
drop_one_value_col = function(df) {
  drop <- NULL
  for (i in 1:dim(df)[2]) {
    if(df |> distinct(df[, i]) |> count() == 1) {
      drop = c(drop, i)
    }
  }
  if (is.null(drop)) {
    return("none")
  }
  else {
    print("Columns dropped:")
    print(colnames(df)[drop])
    df <- df[, -1*drop]
  }
}
strawberry <- strawberry |> drop_one_value_col()
```

```{r}
# separate into 3 tibbles by year
strawberry_2023 = strawberry |> filter(Year == 2023)
strawberry_2021 = strawberry |> filter(Year == 2021)
strawberry_2019 = strawberry |> filter(Year == 2019)
```

Separating composite columns

we want to split `Data.Item` into (fruit, category, item, metric)

```{r}
strawberry <-  strawberry |>
  separate_wider_delim(
    cols = `Data.Item`,
    delim = ",",
    names = c("Fruit", "Category", "Item", "Metric"),
    too_many = "error",
    too_few = "align_start"
  )
## use too_many and too_few to set up the separation operation
```

there is a problem to fix: leading white space

```{r}
strawberry$Category <- str_trim(strawberry$Category, side = "both")
strawberry$Item <- str_trim(strawberry$Item, side = "both")
```

now examine the fruit column – find hidden subcolumns

```{r}
spr <- which(strawberry$Fruit=="STRAWBERRIES - APPLICATIONS")

strw_apps <- strawberry |> slice(spr)

##############################

## this has the census data, too
strw_chem <- strawberry |> slice(-1*spr)
```

now examine the rest of the columns

clean data into different tables, visualize data

my_functions.R -\> put all functions in this.

source("my_functions.R") in my code to have access to those functions.

pick 2 or 3 chemicals to look at. detailed analysis
